Analysis of the 2005--\myrevision{2016} Earthquake Sequence in Northern Iran Using the Visibility Graph Method

### 2017-06-20 results.tex
updates on modifications of the revised version for paag

Next, we examine the relationship between the $b$-values from Table \ref{tab:seismicity} and the $k$-$M$ slopes.

* * *

\myrevision{The $b$-value, by definition, represents the proportion between the amount of large events and that of small events.} Fig.~\ref{fig:regression} shows the scattered results for $k$-$M$ slope versus $b$-value for the three tectonic seismic regions in northern Iran along with the data-points obtained for the analysis of the magnitude-time sequences of the Mexican subduction zone \citep{Telesca2013} and the Pannoninan seismic zone \citep{Telesca2014}, as well as other experimental results \citep{Telesca2014-pone}. \myrevision{According to Scholtz (1968) b-value for microfractures in rocks is reported ranging between 0.11 and 2.58. Using different synthetic and real catalog, in this study we cover a broad range from low b-value (synthetics) to high b-value(subduction zone)}. Fig.~\ref{fig:regression} also includes the results of different linear regressions between the $b$-value and the $k$-$M$ slope. Each regression reflects the addition of new data-points from different studies. Note that the regression improves as new data-points are added, which is indicated by the correlation coefficient $R$, also included in the figure. The correlation fitting various regional seismic data was previously pointed out by \citet{Telesca2014}.

According to these results, a universal relationship between the $b$-value and the $k$-$M$ slope ($m$) can be expressed as:
% 
\begin{equation}
	b = 0.073 + 0.084 m \, .
	\label{eq:universal.bm}
\end{equation}

\begin{figure}[h]%[t]
	\centering
	\includegraphics[width=0.45\textwidth]{figures/pdf/figure-07-rev.pdf} 
	\caption{Correlation between $k$-$M$ slope and $b$-value as drawn from the results of the present study for the region of northern Iranian and three other previous studies, including analysis of the Mexican subduction zone \citep{Telesca2013}, the Pannonian seismic zone \citep{Telesca2014}, and results from two experiments \citep{Telesca2014-pone}. The lines represent linear regressions obtained to fit the different data points, considering different combinations. The values of the correlation coefficient, $R$, are indicated for each regression line. The color version of this figure is available only in the electronic edition.}
	\label{fig:regression}
\end{figure}

Another aspect of interest is the stability of the data points themselves. Note that as presented in Fig.~\ref{fig:regression}, the analysis of the sequences shown in Fig.~\ref{fig:mag-time} only contribute one data point per region of interest. Furthermore, each data-point comes from sequences that vary significantly in terms of the number of events and seismic parameters (see Table \ref{tab:seismicity}).

\citet{Telesca2013} observed that the value of the $k$-$M$ slope is not particularly sensitive to the sample size in the sequence---at least not when considering sufficiently large sequences. On the other hand, as we will see below, if the sequence window is sufficiently small, then the $k$-$M$ slope value shows a relative dependence on time and thus provides insight about the variation of the seismicity as the sequence progresses. 
% \cmmnt{Note also that the threshold value of $M_c$ is significantly smaller for Azerbaijan than for Alborz or Kopeh Dagh. This is due in part to the fact that the latter two zones were more seismically active in the time period under consideration} 
\myrevision{In this study due to using windowing method, we consider a very conservative completeness magnitude}. However, according to \citet{Telesca2012}, the threshold magnitude has a minor effect in the graph parameters.

To further explore the sensitivity of the graph properties to the number of events in each catalog and the value of the minimum magnitude, we randomly picked a significant number of sub-sequences from within the initial catalog compiled for each region, and repeated the analysis for each sub-sequence. In total, for each region, we extracted 200 new sub-sequences from the \myrevision{whole} catalog. The number of events in each sub-sequence was varied randomly but chosen to be large enough to represent the seismic characteristics of each region. In particular, the minimum size of each sub-sequence was set to be $n \geq 200$, and the maximum size in the sequence was set to be as large as the original catalog (\myrevision{whole catalog}; see Table \ref{tab:seismicity}). We forced the random sub-sequences to progress positively in time without altering the natural occurrence of events. In other words, we randomly determined the initial event and the sub-sequence size (number of events to be considered), and then picked that number of events following the initial earthquake in the sub-sequence. Next, we determined the value of $M_c$ and $b$ as previously done for the complete catalogs, created the graph for all events with $M \geq M_c$, and extracted the connectivity degree of the events in each sub-sequence. \myrevision{Note that the results of the whole catalog for each seismic region is different from Fig.~\ref{fig:regression}. In Fig.~\ref{fig:regression} we choose a time dependent completeness magnitude which is conservative and requires inspection. In order to keep uniformity in the random data test we computed the completeness magnitude through GFT for all subsamples and the whole catalog of each seismic region. Determination of completeness magnitude based on the best score of GFT between synthetic catalog and subsamples could be controversial. It is possible to choose very large $M_c$ just because of slightly higher GFT score than smaller magnitudes. This could result in a sequence  of data where is not completely a representative of the domain. However, knowing these source of uncertainties, the randomly generated data conforms the fact that the method is fairly robust for the catalog size and completeness magnitude; where we can easily distinguish the 3 seismic region border. There are some outlier data due to subsamples of aftershocks of big earthquakes. These sequences represent different $k-M$ and $b-value$ relationship. These sequences do not represent the dominant seismicity pattern of the region. In the declustered catalog (is not presented here) we do not see these outliers.}

@@

Fig.~\ref{fig:random}a shows scattered points for all the individual $k$-$M$ slope and $b$-value pairs for all the sub-sequences that were randomly picked from the initial catalogs of each region in northern Iran (small symbols). Figures \ref{fig:random}b and \ref{fig:random}c show the variability of the $k$-$M$ slope and $b$-value, respectively. These latter figures show the mean values of each parameter for all the random sub-sequences and the amplitudes of $\pm 1$ standard deviation, as well as the coefficients of variation (in percentage). Figure \ref{fig:random}a also includes the data points corresponding to the complete catalogs (large solid symbols); the points corresponding to the mean values of the $k$-$M$ slope and $b$, from Figs.~\ref{fig:random}b and \ref{fig:random}c (large hollow symbols); and the universal linear regression from equation (\ref{eq:universal.bm}) (thick line), as well as the linear regressions for northern Iran when using the complete catalogs (dashed thin line) and the random sub-catalogs (continuous thin line).

The comparison between the mean values of the random sequences analysis versus the complete catalogs presented in Fig.~\ref{fig:random} indicates that there exists only a small bias, which is well within the standard deviation of the different value samples. We note that the values of the $k$-$M$ slope are slightly smaller when obtained with the random sequence analysis, whereas the $b$ values seem more stable.

\begin{figure*}%[t]
	\centering
	\includegraphics[width=0.75\textwidth]{figures/pdf/figure-09-rev.pdf} 
	\caption{Variation of <$T_c$>, $k$-$M$ slope and $b$-value as with respect to time for three seismic zones for a moving window analysis of the visibility graphs of sub-sequences of \myrevision{50} consecutive events, along with the magnitude of events in the 2005--\myrevision{2016} period. Black triangles indicate the occurrence of major earthquakes in each of the regions, with the corresponding magnitude at the top of each symbol. In the middle frame of each region, the dashed line corresponds to the $k$-$M$ values, whereas the continuous plot shows the variation of the $b$-value. The color version of this figure is available only in the electronic edition.}
	\label{fig:tc}
\end{figure*}

@@

>>>>

Next, we examine the relationship between the $b$ values from Table \ref{tab:seismicity} and the $k$-$M$ slopes. \myrevision{Figures \ref{fig:regression}a and \ref{fig:regression}a show the scattered results of the $k$-$M$ slope versus $b$-value, for the complete and declustered catalogs, respectively. Along with the data-points obtained for our analysis of the three seismic regions of northern Iran, in each plot, we also include the data points corresponding to the magnitude-time sequences of the Mexican subduction zone provided by \citet{Telesca2013}, the Pannoninan seismic zone from \citet{Telesca2014}, the experimental results obtained by \citet{Telesca2014-pone}, and the result from a similar preliminary analysis done by \citet{Azizzadeh_2017_SSA} for California. These figures also include the linear trends of the data points, as they are aggregated (i.e., each regression reflects the addition of new data-points from different studies). In the figure, we indicate which data points were considered for each of the computed linear regressions and their corresponding correlation coefficient, $R$.}

\myrevision{Here, as well as in other figures, we present our results for both the complete and declustered catalogs to shed light upon the local differences for the case of northern Iran, and to contribute, globally, to the discussion of the appropriate treatment and selection of catalogs for the use of the visibility graph method. We note, for instance, that the data points in Fig.~\ref{fig:regression} from \citet{Telesca2012} and \citet{Telesca2013} used whole catalogs, whereas those from \citet{Telesca2014} and \citet{Telesca2016} used declustered catalogs.}

\myrevision{Note that, overall, the linear trends shown in Fig.~\ref{fig:regression} hold independently of the differences in the studies in terms of number of events, seismic characteristics, and the treatment of the catalogs. Also note that these trends tend to improve with increasing number of data points, as reflected by the increasing value of $R$. Regarding this, we should note that the data points in Fig.~\ref{fig:regression} cover what would otherwise be considered a large span of $b$ values. While we recognize that most earthquake data lead to $b$ values ranging between 0.8 and 1.2, it is also known that there are regions with low- and high-$b$ values \citep[e.g.,][]{Singh_1983_BSSA, Pacheco_1992_N, Nakaya_2006_JGR}. In addition, \citet{Scholz_1968_BSSA} reported that experimental microfractures in rocks, similar to those observed in earthquakes, exhibit $b$ values ranging between 0.11 and 2.58. Considering that the $b$ value, by definition, represents the proportion between the amount of large and small events in a region, the trends shown in Fig.~\ref{fig:regression} contribute to the suggestion made by \citet{Telesca2014} about the universal character of the relationship between $b$ and the $k$-$M$ slope.}

According to these results, a universal relationship between $b$ and the $k$-$M$ slope ($m$) can be expressed as
% 
\begin{equation}
\myrevision{
	b = 0.073 + 0.084 m
	\label{eq:universal.bm}
}
\end{equation}
%
\myrevision{for the case in which we consider the complete catalogs for northern Iran, and}
% 
\begin{equation}
\myrevision{
	b = 0.077 + 0.085 m
	\label{eq:universal.bm.dc}
}
\end{equation}
%
\myrevision{for the case in which we consider the declustered catalogs for northern Iran. This indicates that once combined with the data points of other regions, the option of considering either catalog leads to very similar results as, on average, the data points oscillate about the same universal relationship. Locally, however, for the case of northern Iran trends alone, it seems that the declustered catalog leads to a closer fit with the universal relationship. Separately (not shown here for brevity), we also observed that, for the particular sequences under study, the results seem to be more sensitive to less conservative values of $M_c$, as the choice of a smaller $M_c$ leads to larger datasets. \citet{Telesca2012} and \citet{Telesca2013}, however, pointed out that for sufficiently long sequences, the threshold magnitude have only small effects on the graph parameters.}

\begin{figure*}[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-07-rev.pdf} 
	\caption{Correlation between $k$-$M$ slope and $b$-value as drawn from the results of the present study for the region of northern Iranian and three other previous studies, including analysis of the Mexican subduction zone \citep{Telesca2013}, the Pannonian seismic zone \citep{Telesca2014}, and results from two experiments \citep{Telesca2014-pone}. The lines represent linear regressions obtained to fit the different data points, considering different combinations. The values of the correlation coefficient, $R$, are indicated for each regression line. The color version of this figure is available only in the electronic edition.}
	\label{fig:regression}
\end{figure*}

To further explore the sensitivity of the graph properties to the number of events in each catalog and the value of the minimum magnitude, we randomly picked a significant number of sub-sequences from within the initial catalog compiled for each region, and repeated the analysis for each sub-sequence. In total, for each region, we extracted 200 new sub-sequences from the \myrevision{whole} catalog. The number of events in each sub-sequence, $n$, was varied randomly but chosen to be large enough to represent the seismic characteristics of each region. In particular, the minimum size of each sub-sequence was set to be $n \geq 200$, and the maximum size in the sequence was set to be as large as the original \myrevision{whole} catalog (see Table \ref{tab:seismicity}). We forced the random sub-sequences to progress positively in time without altering the natural occurrence of events. In other words, we randomly determined the initial event and the sub-sequence size (number of events to be considered), and then picked that number of events following the initial earthquake in the sub-sequence. Next, we determined the value of $M_c$ and $b$ \myrevision{to obtain the complete and declustered catalogs for each region}, created the graph for all events with $M \geq M_c$, and extracted the connectivity degree of the events in each sub-sequence. \myrevision{Because this needed to be done for a large number of sub-sequences, in this part of the analysis we used the goodness-of-fit test (GFT) method to determine $M_c$ systematically, instead of the MAXC method. The GFT method follows the recommendations of \citet{Wiemer2000} and \citet{Wiemer2001}, and leads to less conservative values of $M_c$, which is preferable to keep the subsequences large enough.}

@@

\myrevision{Fig.~\ref{fig:random} shows the results for this analysis on random subsequences for both the complete and declustered catalogs. In particular, Figs.~\ref{fig:random}a and \ref{fig:random}d show} scattered points for all the individual $k$-$M$ slope and $b$-value pairs for all the subsequences that were randomly picked from the \myrevision{complete and declustered} catalogs of each region in northern Iran (small symbols)\myrevision{, respectively}. Figures \ref{fig:random}b \myrevision{and \ref{fig:random}e show the variability of the $k$-$M$ slope, and Figs.~\ref{fig:random}c and \ref{fig:random}f show the variability of $b$. The variability plots} show the mean values of each parameter for all the random subsequences and the amplitudes of $\pm 1$ standard deviation, as well as the coefficients of variation (in percentage). \myrevision{Figures \ref{fig:random}a and \ref{fig:random}d also include the data points corresponding to the regional analysis of the catalogs, the points corresponding to the mean values of the $k$-$M$ slope and $b$, and the universal linear regression from equations (\ref{eq:universal.bm}) and (\ref{eq:universal.bm.dc}), as well as the linear regressions for northern Iran when using the full-sequence catalogs (dashed thin line) and the random subsequence catalogs (continuous thin line).}

\myrevision{In essence, the results in Fig.~\ref{fig:random} show that, despite the variability in the random subsequences and the differences in the computation of $M_c$, both of which are sources of additional uncertainty, the analysis produces an outcome in fair agreement with the universal relationships. Nonetheless, the analysis of the random subsequences in the complete catalogs produces some outliers, whereas the processing of the declustered subsequences tends to concentrate more evenly around the universal trend. Note that, likely due to the removal of foreshocks and aftershocks, the declustered random subsequences exhibit a clearer distinction between the three seismic zones.}

\myrevision{While these sequences do not represent the dominant seismicity pattern of the region, the comparison between the mean values of the random sequences analysis versus the complete catalogs (large, empty and solid symbols in Fig.~\ref{fig:random}) indicates that there exists only a small bias, which is well within the standard deviation of the different value samples. Note also that the values of the $k$-$M$ slope in Fig.~\ref{fig:random} are slightly smaller when obtained with the random sequence analysis than when done for the regional sequences, whereas the $b$ values seem more stable. This explains why the regressions of the random subsequences are less steep than the universal regressions, in both cases. In a separate initial analysis, not included here for brevity, we observed that lower, less conservative selections of $M_c$ can lead to random subsequences that yield regressions in better agreement with the universal trend. Using lower values of $M_c$, however, may be misleading. In this sense, larger catalogs in better instrumented regions may shed light on the stability of the $m$-$b$ relationship, which is the reason we included the data point from \citet{Azizzadeh_2017_SSA} for southern California in Fig.~\ref{fig:regression}, which falls well in line with the universal regression.}

* * * HERE * * *

\begin{figure*}%[t]
	\centering
	\includegraphics[width=0.75\textwidth]{figures/pdf/figure-09-rev.pdf} 
	\caption{Variation of $<$$T_c$$>$, $k$-$M$ slope and $b$-value as with respect to time for three seismic zones for a moving window analysis of the visibility graphs of sub-sequences of \myrevision{50} consecutive events, along with the magnitude of events in the 2005--\myrevision{2016} period. Black triangles indicate the occurrence of major earthquakes in each of the regions, with the corresponding magnitude at the top of each symbol. In the middle frame of each region, the dashed line corresponds to the $k$-$M$ values, whereas the continuous plot shows the variation of the $b$-value. The color version of this figure is available only in the electronic edition.}
	\label{fig:tc}
\end{figure*}

@@

% OLD MATERIAL
% -----------------------------------------------------------------

% Another aspect of interest is the stability of the data points themselves. Note that as presented in Fig.~\ref{fig:regression}, the analysis of the sequences shown in Fig.~\ref{fig:mag-time} only contribute one data point per region of interest. Furthermore, each data-point comes from sequences that vary significantly in terms of the number of events and seismic parameters (see Table \ref{tab:seismicity}).

% \myrevision{According to Scholtz (1968) b-value for microfractures in rocks is reported ranging between 0.11 and 2.58. Using different synthetic and real catalog, in this study we cover a broad range from low b-value (synthetics) to high b-value(subduction zone)}.

% Fig.~\ref{fig:regression} also includes the results of different linear regressions between the $b$-value and the $k$-$M$ slope. Each regression reflects the addition of new data-points from different studies. Note that the regression improves as new data-points are added, which is indicated by the correlation coefficient $R$, also included in the figure. The correlation fitting various regional seismic data was previously pointed out by \citet{Telesca2014}.

% \cmmnt{Note also that the threshold value of $M_c$ is significantly smaller for Azerbaijan than for Alborz or Kopeh Dagh. This is due in part to the fact that the latter two zones were more seismically active in the time period under consideration} 
% \myrevision{In this study due to using windowing method, we consider a very conservative completeness magnitude}. However, according to \citet{Telesca2012}, the threshold magnitude has a minor effect in the graph parameters.

% Moving on, \citet{Telesca2013} observed that the value of the $k$-$M$ slope was not particularly sensitive to the sample size in the sequence---at least not when considering sufficiently large sequences. On the other hand, as we will see below, if the sequence window is sufficiently small, then the $k$-$M$ slope value shows a relative dependence on time and thus provides insight about the variation of the seismicity as the sequence progresses.

% Note that the results of the whole catalog for each seismic region is different from Fig.~\ref{fig:regression}. In Fig.~\ref{fig:regression} we choose a time dependent completeness magnitude which is conservative and requires inspection. In order to keep uniformity in the random data test we computed the completeness magnitude through GFT for all subsamples and the whole catalog of each seismic region. Determination of completeness magnitude based on the best score of GFT between synthetic catalog and subsamples could be controversial. It is possible to choose very large $M_c$ just because of slightly higher GFT score than smaller magnitudes. This could result in a sequence of data where is not completely a representative of the domain. However, knowing these source of uncertainties, the randomly generated data conforms the fact that the method is fairly robust for the catalog size and completeness magnitude; where we can easily distinguish the 3 seismic region border. There are some outlier data due to subsamples of aftershocks of big earthquakes. These sequences represent different $k-M$ and $b-value$ relationship. These sequences do not represent the dominant seismicity pattern of the region. In the declustered catalog (is not presented here) we do not see these outliers.}
